### 2015-05-25 implementation.tex
Theory and shit

\subsection{OpenVPN}
% By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.

\lstset{language=c}
\begin{lstlisting}[caption=openvpn compress then encrypt -- sample from \texttt{forward.c}, label=list:openvpn-workflow]
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

@@

\begin{figure}[ht]
\includegraphics[width=\textwidth]{OpenVPN-packetflow}
\caption{OpenVPN packet flow as advertized on the openVPN wiki.}{}
\label{fig:openvpn-packet-flow}
\end{figure}

@@

\noindent The second liberty has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\cite{chap:openssh-patch}.

>>>>

\subsection{OpenVPN}
% By default, point-to-point, but since the version 2.0 can be configured in server mode, hence being able to manage hundreds of clients.
% https://openvpn.net/index.php/open-source/339-why-ssl-vpn.html
% https://openvpn.net/index.php/open-source/335-why-openvpn.html
% https://openvpn.net/papers/BLUG-talk/index.html

OpenVPN is also used by the Dutch government to secure all its communications~\cite{openvpn-nl}.

Uses regular TCP/UDP network protocols, which can be an advantage over IPSec if the ISP decides to block the latter.

OpenVPN offers two different TUN/TAP virtual interfaces:
\begin{description}
	\item[TUN] Virtual layer 3 IP tunnel, seen as a point-to-point interface from the kernel point of view.
	\item[TAP] Virtual layer 2 ethernet tunnel.
\end{description}

Since OpenVPN aims at network security at the application level, it needs virtual network interface to which to push and from which to read the secured data.
The \texttt{tun} adapter will be used when OpenVPN wants to establish a point-to-point connection between two client, while the \texttt{tap} device is used when there is one server acting as a bridge, hence managing lots of clients at the same time.

OpenVPN as two choices when it wants to sends its data over the internet: either encapsulating its IP packet into TCP or UDP packets.
The problem of encapsulating IP into TCP is that inside IP packets are already TCP frames.
The reason is that IP has been designed to work on unreliable networks, and TCP includes in its standard protocols to retry and drop packets.
Hence, encapsulating an IP packet into a TCP packet produces a redundancy by nesting reliability layers.
UDP is the unreliable counterparty of TCP, offering a better alternative for encapsulation.
If we refer to the figure~\ref{fig:openvpn-packet-flow}, \texttt{iperf} would send its regular \texttt{IP(TCP)} packet over the virtual \texttt{tun} device, which will then compress, fragment and encrypt the frame before encaspulating it into a new UDP packet that will finally be sent through the physical \texttt{eth0} interface.

\begin{figure}[ht]
\includegraphics[width=\textwidth]{OpenVPN-packetflow}
\caption{OpenVPN packet flow}{it shows two hosts using \texttt{iperf} inside an OpenVPN tunnel. This figure comes from the OpenVPN documentation~\cite{openvpn-doc-workflow} and has inverted the encryption and fragmentation.}
\label{fig:openvpn-packet-flow}
\end{figure}

\noindent At this point, it is important to note that if the figure~\ref{fig:openvpn-packet-flow} shows a workflow in which the encryption taking place before the fragmentation, a quick look at the source code proves the opposite.
A sample of the said code has been writen down in the listing~\ref{list:openvpn-workflow}.

\lstset{language=c}
\begin{lstlisting}[caption={openvpn compress then encrypt -- sample from \texttt{forward.c}. It clearly shows that the order of operations in the packet workflow is compression, then fragmentation and finally encryption.}, label=list:openvpn-workflow, float]% The 'float' option makes the listing unbreakable.
/* Compress, fragment, encrypt and HMAC-sign an outgoing packet. */
void encrypt_sign (struct context *c, bool comp_frag)
{
	struct context_buffers *b = c->c2.buffers;
	const uint8_t *orig_buf = c->c2.buf.data;

@@

An other way to look at this workflow is to completely isolate the OpenVPN part from both hosts, and exporting it to an independent device such as a router.
The hosts would be stripped off the virtual interface -- \texttt{tun0} would become physical, and would communicate normally over the internet, at least from their point of view.
All their communications would be routed by the router of their local area network, incomming on a physical \texttt{tun0} interface and outputed to the internet on the \texttt{eth0} interface.
Such implementation are used when an adminstrator wishes to secure a whole local area network without bloating each client with a VPN application.
In this case, there exist router firmware packages running OpenVPN, such as DD-WRT and OpenWRT.
Whatever the method used, the crucial point is that the VPN tunnel usage is transparent to the application.\newline{}

OpenVPN relies on a cryptographic library, OpenSSL in our case, for all the security operations: encryption, signature, certificate management, etc.

\noindent Note that if one wants to use the \texttt{none} cipher with a version of OpenVPN ealier than 2.3.7, he would have to update it using a community patch~\cite{openvpn-patch-none}.

@@

\noindent The second liberty has been taken to prevent the user to strip himself from data authenticity and integrity.
However, in the case of testing and benchmarking, leaving the MAC behind can be interresting, especially if it is not offloaded in hardware such as the encryption in our case.
In order to add this feature to OpenSSH, we wrote a patch to apply on OpenSSH 6.7, available in appendix~\ref{chap:openssh-patch}.

### 2015-05-25 technical-background.tex
Theory and shit

\section{Operating system}

@@

\section{FPGA}

@@

\section{Cryptography}

@@

\subsubsection{SHA}
Keyed signature algorithm, several versions in place. SHA-1 is depreciated, SHA-2 is widely used and SHA-3 is already defined and begins to be implemented.

@@

\section{Network and VPN implementation}

@@

There exist several major implementations of VPN: SSL, IPSec, L2TP and PPTP.
The later was developped by a vendor consortium and proposed in the RFC 2637 and will not be discussed further.

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.

\subsection{IPSec}
% IPSec as a protocol, strongswan come in the 'implementation' chapter.

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Operating system}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{FPGA}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Cryptography}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

Cryptography is the corner stone of security.
The four main goals are the following, as defined in~\cite{Menezes1996}:
\begin{description}
	\item[Confidentiality] keeping information secret from all but those who are authorized to see it.
	\item[Integrity] ensuring information has not been altered by unauthorized or unknown means.
	\item[Source Authentication] corroborating the source of information.
	\item[Non-repudiation] preventing the denial of previous commitments or actions.
\end{description}

In order to achieve those, four cryptographic primitives are needed: symmetric and asymmetric ciphers, message digests and digital signatures.

@@

\begin{figure}
\includegraphics[width=\textwidth]{nist-cbc}
\caption{CBC encryption and decryption diagram}{taken from the NIST recommendation~\cite{nist-sp800-38A}.}
\label{fig:cbc-encrypt-decrypt}
\end{figure}

\begin{figure}
\includegraphics[width=\textwidth]{nist-gcm-encrypt}
\caption{GCM encryption diagram}{taken from the NIST specification~\cite{mcgrew2005}. The ciphertext blocks are formed by \textit{xor}-ing the encrypted counter and the plaintext. The tag is generated by a chain of ciphertext \textit{xor}-ing with Galois field multiplicated data. The decryption works excatly the same way, except the plaintext and ciphertext are swapped.}
\label{fig:gcm-encrypt}
\end{figure}

\subsection{Message digest}
%Not the same as digital signature because we use the same key for both MAC values.
A message digest is the result of a one-way mathematical function of a fixed size.
Those hash functions are of two types~\cite{infof405}: manipulation detection codes (MDC) to guarantee integrity and message authentication codes (MAC) to guarantee both integrity and source authentication.

An MDC $h(x)$ can follow an iterative construction for a message $x$ including $t$ blocks:
\[
\begin{dcases}
	H_0 = \mbox{initial value}\\
	H_i = f(H_{i-1}, x_i), \mbox{with } i \in [1,t]\\
	h(x) = H_t\\
\end{dcases}
\]

Based on this design and adding a key to the process, the RFC 2104~\cite{rfc2104} defines a MAC:
\[
	HMAC(k, x) = h((k\oplus opad)|h((k\oplus ipad)|x))
\]
with a key $k$, and two padding block added for security concerns: an outer pad $opad$ and an inner pad $ipad$.

\noindent There exist a wide varety of MDCs, ranging from block cipher based such as Miyaguchi-Preneel, customized such as MD5, SHA-1 and SHA-2, or built using modular arithmetic such as MASH-1.

In both schemes, data integrity can be guaranteed because the flip of one bit will irremediably change the digest.
However, only a MAC can ensure source authentication since it is the only one based on a shared secret key.

Now rises the question of what and when authenticating.
\citet{Bellare2000} prooved that the most secure solution is to encrypt then compute the MAC from the ciphertext.
We will see in section~\ref{sec:theory-network} that if IPSec follows this recommandation, SSL/TLS does not and MAC first the plaintext then encrypt the message.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Network and VPN implementation}\label{sec:theory-network}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

There exist several major implementations of VPN: SSL, IPSec, L2TP and PPTP.
The later was developped by a vendor consortium leaded by Microsoft and proposed in the RFC 2637 and will not be discussed further.

\subsection{SSL/TLS}
% Introduce SSL/TLS, talk about the protocol, the key exchange and stuff, but leaver OpenVPN for the 'implementation' chapter.
% Question the security? Apparently, it does mac-then-encrypt, which is insecure regarding certain types of attacks \cite{cryptoeprint:2001:045}.
% Plus, SSLv3 is to be deprecated, according to a queued RFC: http://www.rfc-editor.org/internet-drafts/draft-ietf-tls-sslv3-diediedie-03.txt, and is not supported anymore by many servers since poodle.
Application level security.

\subsection{IPSec}
% IPSec as a protocol, strongswan come in the 'implementation' chapter.
Modification of the IP stack in the kernel space.
IPSec is a network level security; it examines incomming IP packets and checks if there exists a security association with the destination, and decrypt it on-the-fly if necessary.

One of the main disadvantage compared to a user-space VPN is the difficulty to traverse NAT.
